---
title: "Assignment 5: Data Visualization"
author: "Student Name"
output: pdf_document
geometry: margin=2.54cm
---
{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/path/to/folder/")
```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.

The completed exercise is due on Friday, Oct 14th @ 5:00pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse, 
lubridate, & cowplot packages. Upload the NTL-LTER processed data files for  
nutrients and chemistry/physics for Peter and Paul Lakes (use the tidy 
[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`] version) and 
the processed data file for the Niwot Ridge litter dataset (use the
[`NEON_NIWO_Litter_mass_trap_Processed.csv`] version). 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r}
#1 
library(tidyverse)
library(lubridate)
library(dplyr)
library(cowplot)
getwd()
setwd('c:/Users/Zhiteng Ma/Desktop/EDA-Fall2022-main/Data/Processed/')
NTL_LTER_LAKE_PeterPaul <-read.csv('NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv',stringsAsFactors =TRUE)
NEON_NIWO <- read.csv('NEON_NIWO_Litter_mass_trap_Processed.csv', stringsAsFactors = TRUE)
#2 
NTL_LTER_LAKE_PeterPaul$sampledate <- as.Date(NTL_LTER_LAKE_PeterPaul$sampledate, format = "%Y-%m-%d")
NEON_NIWO$collectDate <- as.Date(NEON_NIWO$collectDate, format = "%Y-%m-%d")
```
## Define your theme
3. Build a theme and set it as your default theme.

```{r}
#3
theme_default <- theme_set(theme_bw())
theme_set(theme_default)
#theme_update(panel.grid.minor = element_line(colour = "red"))

```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best 
practices for data visualization. Ensure your theme, color palettes, axes, 
and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with 
separate aesthetics for Peter and Paul lakes. Add a line of best fit and 
color it black. Adjust your axes to hide extreme values (hint: change the 
limits using `xlim()` and/or `ylim()`).

```{r, fig.height = 3, fig.width = 4, warning = FALSE}
#PeterPaul <- 

summary(NTL_LTER_LAKE_PeterPaul$lakename)
NTL_LTER_LAKE_PeterPaul1 <- filter(NTL_LTER_LAKE_PeterPaul, lakename == "Paul Lake")

NTL_LTER_LAKE_PeterPaul2 <- filter(NTL_LTER_LAKE_PeterPaul, lakename == "Peter Lake")

NTL_LTER_LAKE_PeterPaul3 <- ggplot(NULL, aes(x = tp_ug, y = po4)) +
  geom_point(data=NTL_LTER_LAKE_PeterPaul1, color = "blue") +
  geom_point(data=NTL_LTER_LAKE_PeterPaul2) +
  geom_smooth(data=NTL_LTER_LAKE_PeterPaul1, color = "black") +
  xlim( 0 , 140 ) +
  ylim( 0 , 40 )
print(NTL_LTER_LAKE_PeterPaul3)

#4

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and 
(c) TN, with month as the x axis and lake as a color aesthetic. Then, create 
a cowplot that combines the three graphs. Make sure that only one legend is 
present and that graph axes are aligned. 

Tip: R has a build in variable called `month.abb` that returns a list of months;
see <https://r-lang.com/month-abb-in-r-with-example>

```{r}
#5

Tem <-
  ggplot(NTL_LTER_LAKE_PeterPaul, aes(x = month.abb[month], y = temperature_C)) +
  geom_boxplot(aes(color = lakename))
print(Tem)

TP_ug <-
  ggplot(NTL_LTER_LAKE_PeterPaul, aes(x = month.abb[month], y = tp_ug)) +
  geom_boxplot(aes(color = lakename))
print(TP_ug)

TN_ug  <-
  ggplot(NTL_LTER_LAKE_PeterPaul, aes(x = month.abb[month], y = tn_ug)) +
  geom_boxplot(aes(color = lakename))
print(TN_ug)

plot_grid(Tem, TP_ug, TN_ug, labels=c("A", "B", "C"), ncol = 2, nrow = 2)

```

Question: What do you observe about the variables of interest over seasons 
and between lakes? 

> Answer: 

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the 
"Needles" functional group. Plot the dry mass of needle litter by date and 
separate by NLCD class with a color aesthetic. (no need to adjust the name 
of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into 
three facets rather than separated by color. 

```{r}
#6

#7


```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: